---
title: "AF1204 Portfolio — ESG & Valuation result tables (R cross-validation)"
output: html_notebook
---

Adapted from the course `WK10_ResultTables_Moodle.Rmd` (Week 10). Purpose here:
reproduce the Python/statsmodels Models 1–4 in R and export publication-style
stargazer tables to `../data/regression_r.html`, so the deployed site can embed
them and the README can report Python↔R agreement on the ESGscore coefficient.

Run it in the course Codespace or RStudio (R is not available in every environment):
`rmarkdown::render("analysis_R/WK10_ResultTables_AF1204.Rmd")` or press *Run All*.

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(modelsummary)
library(correlation)
library(stargazer)
```

## Load the provided data (course files)

```{r load}
ESGscore_TobinsQ <- read.csv("../course_materials/week10/notebooks/public/ESGscore_TobinsQ_Moodle.csv")
load("../course_materials/week10/notebooks/public/ESG_data_Moodle.RData")  # -> ESG_data
```

## Variable construction (identical to the course Rmd)

```{r construct}
ESG_data_new <- ESG_data |>
  mutate(date = as.character(date)) |>
  left_join(ESGscore_TobinsQ, by = c("instrument" = "tic", "date" = "datadate")) |>
  mutate(
    at_pos  = ifelse(at > 0, at, NA),
    ceq_pos = ifelse(ceq > 0, ceq, NA),
    q         = (prcc_f * csho + dltt + dlc) / at_pos,  # Tobin's Q
    ESGscore  = esg_metric,
    Firm_Size = log(at),
    Leverage  = (dltt + dlc) / ceq_pos,
    Year      = fyear,
    Industry  = sich
  ) |>
  select(source, q, ESGscore, Firm_Size, Leverage, Year, Industry) |>
  drop_na()

ESG_data_B <- ESG_data_new |> filter(source == "Provider_B") |> select(-source)
ESG_data_A <- ESG_data_new |> filter(source == "Provider_A") |> select(-source)

rm(ESG_data_new, ESG_data, ESGscore_TobinsQ)  # remove data no longer needed
gc()  # garbage collection to free up memory
```

## Descriptive statistics & correlations

```{r describe}
datasummary(q + ESGscore + Firm_Size + Leverage ~
              (N = N) + (Mean = Mean) + (Median = Median) +
              (SD = SD) + (Min = Min) + (Max = Max),
            data = ESG_data_B, output = "default")

ESG_data_B |>
  select(q, ESGscore, Firm_Size, Leverage) |>
  correlation() |>
  datasummary_correlation(stars = TRUE)
```

## Models 1–4 (same specification as `pipelines/pipeline_4_merge.py`)

```{r models}
model1 <- lm(q ~ ESGscore, data = ESG_data_B)
model2 <- lm(q ~ ESGscore + Firm_Size + Leverage, data = ESG_data_B)
model3 <- lm(q ~ ESGscore + Firm_Size + Leverage + factor(Industry), data = ESG_data_B)
model4 <- lm(q ~ ESGscore + Firm_Size + Leverage + factor(Industry) + factor(Year),
             data = ESG_data_B)
```

## Main table → HTML for the deployed site

```{r main-table}
suppressWarnings(
  stargazer(model1, model2, model3, model4,
            type = "html", out = "../data/regression_r.html",
            title = "ESG score and firm valuation (Tobin's q) — estimated in R",
            dep.var.labels = "q",
            column.labels = c("Model 1", "Model 2", "Model 3", "Model 4"),
            covariate.labels = c("ESGscore", "Firm Size", "Leverage"),
            keep = c("ESGscore", "Firm_Size", "Leverage"),
            star.cutoffs = c(0.1, 0.05, 0.01),
            notes = "Standard errors in parentheses; ***p<0.01, **p<0.05, *p<0.1",
            notes.append = FALSE,
            add.lines = list(c("Industry Indicators", "No", "No", "Yes", "Yes"),
                             c("Year Indicators", "No", "No", "No", "Yes")),
            keep.stat = c("n", "rsq"),
            digits = 3, no.space = TRUE, align = TRUE,
            font.size = "small", single.row = FALSE)
)
```

## Robustness (alt provider, large firms, 2010–2019) — appended to the same HTML

```{r robustness}
m_main  <- model4
m_alt   <- lm(q ~ ESGscore + Firm_Size + Leverage + factor(Industry) + factor(Year),
              data = ESG_data_A)
m_large <- lm(q ~ ESGscore + Firm_Size + Leverage + factor(Industry) + factor(Year),
              data = ESG_data_B |> filter(Firm_Size > median(Firm_Size)))
m_1019  <- lm(q ~ ESGscore + Firm_Size + Leverage + factor(Industry) + factor(Year),
              data = ESG_data_B |> filter(Year >= 2010 & Year <= 2019))

robust_html <- suppressWarnings(
  capture.output(
    stargazer(m_main, m_alt, m_large, m_1019,
              type = "html",
              title = "Robustness analysis — estimated in R",
              dep.var.labels = "q",
              column.labels = c("Main Results", "Alt. ESG Score", "Large Firms", "2010-2019"),
              keep = "^ESGscore", covariate.labels = "ESGscore",
              star.cutoffs = c(0.1, 0.05, 0.01),
              notes = "Standard errors in parentheses; ***p<0.01, **p<0.05, *p<0.1",
              notes.append = FALSE,
              add.lines = list(c("Controls", "Yes", "Yes", "Yes", "Yes"),
                               c("Industry Indicators", "Yes", "Yes", "Yes", "Yes"),
                               c("Year Indicators", "Yes", "Yes", "Yes", "Yes")),
              keep.stat = c("n", "rsq"),
              digits = 3, no.space = TRUE, align = TRUE,
              single.row = FALSE, font.size = "small")
  )
)
write(robust_html, file = "../data/regression_r.html", append = TRUE)
```

## Python ↔ R agreement check

The Python run (`data/regression_py.json`) gives Model 4: **ESGscore = 0.0026 (se 0.0012), N = 7,895**.
The chunk below prints the R equivalent — the two should match to 4 decimal places
because the sample construction and specification are identical.

```{r agreement}
co <- summary(model4)$coefficients["ESGscore", ]
cat(sprintf("R Model 4 ESGscore: coef = %.4f, se = %.4f, N = %d\n",
            co["Estimate"], co["Std. Error"], nobs(model4)))
```
